\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-12-04 tex/gitlinks.tex
more text; features

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
more text; features

@@

\section{Motivation}
\label{sec:motivation}
See Figure~\ref{fig:nullspace_ensemble}.

@@

>>>>

\features{open-source figures \codeicon; equation unit tests \testpassicon}

@@

\section{Motivation}
\label{sec:motivation}

Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
as a function of spherical harmonic degree for three different
(purely hypothetical) scenarios: a single high signal-to-noise ratio (SNR)
observation of a stellar light curve at a random
observer inclination (thin blue curves), a set of three observations
of the same star at three \emph{different} random inclinations (thin
orange curves), and a set of ten observations at ten different
random inclinations (thin green curves).
The denser, opaque curves show the average for each scenario over 300 trials.
The posterior shrinkage $R$ for a given
spherical harmonic coefficient is computed from
%
\begin{align}
    R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
    \frac{\sigma^2}{\sigma_0^2}
\end{align}
%
where $\sigma_0^2$ is the prior variance
and $\sigma^2$ is the posterior variance.
It is a measure of how informative a measurement is about a
given mode on the surface in the limit of infinite SNR
and is independent of what the stellar surface actually looks like.
If, at infinite SNR and with a completely uninformative prior,
a particular mode can be learned exactly from a dataset, the posterior
shrinkage is defined to be unity. Conversely, if the data is completely
unconstraining, $R$ will tend to zero.

@@

Let us first consider the blue curves, in which a light curve is
collected from a star viewed at a single random inclination (which, of course, is
all we can do from our fixed vantage point on Earth).
For some of the low-degree modes, the shrinkage is relatively high: it is
fairly easy to constrain the dipole moment from a light curve, as this is
usually the dominant sinusoidal signal. However, as the degree $l$
increases, the shrinkage decreases dramatically: at $l = 14$, corresponding
to features on scales of roughly $15^\circ$, the light curve
can only tell us about $\sim 10\%$ of the total information about what the
surface looks like. As $l$ increases further, $R$ tends to zero.
Another important feature of the shrinkage is that it is exactly zero for
all odd-degree modes above $l = 1$. This is a well-known fact: all odd spherical
harmonics other than the dipole are in the null space \emph{regardless of
    inclination} \citep[e.g.,][]{Luger2019}. In other words, these spherical
harmonics are perfectly antisymmetric in projection over the unit disk
when viewed from any orientation. Absent structure to break these symmetries
(see \S\ref{sec:limbdark}), we simply cannot learn anything about these modes from
stellar light curves. If we average over the shrinkage for all modes up to $l=15$,
we find that a single light curve measurement can only tell us $\sim 9\%$
of the information about the surface.

The orange curves correspond to the scenario in which we can collect light curves
from a star viewed at three different random inclinations. This is obviously
impossible in practice, but it is still instructive as a thought experiment.
Interestingly, the shrinkage increased at all even spherical harmonic degrees
(the odd degrees, as we mentioned above, are always invisible).
Note that since we are in the limit of infinite SNR, the fact that we have
three times the data as we did in the previous scenario is irrelevant: the
increase in the shrinkage is instead due to the fact that our observations
from different vantage points broke some degeneracies in the problem.
This is a consequence of the fact that the null space (for the even modes)
is a strong function of the inclination: modes that are invisbile
at one inclination can actually project into the flux at a different inclination.
If we average over all modes, we obtain a total shrinkage of $\sim 24\%$
for $l\leq15$ in this case.

Finally, the green curves correspond to observations taken at ten random
inclinations. In this scenario, the posterior shrinkage is nearly $100\%$
for all even modes (it asymptotically approaches $100\%$ as the number
of light curves increases further). The total shrinkage is $\sim 47\%$
and approaches $50\%$ as both $l_\mathrm{max}$  and the number of light curves
tend to infinity. Thus, if we were able to measure a light curve of a star
from many different inclinations, the null space would consist \emph{only}
of the odd modes, and our data would tell us exactly half of all the
information about the surface.

As we mentioned above, it is obviously not possible to observe a \emph{single}
star from many different inclinations (at least not at present!).
However, it is certainly the case that we can observe \emph{many} stars
at a wide range of inclinations. Those stars certainly have different
surface intensity distributions, but what if we restricted ourselves to
light curves of ``similar'' stars?
